---
title: "Chicago Crime"
author: "Rebecca Oh"
output: html_document
runtime: shiny
---

```{r load_packages, include=FALSE}
library(tidyverse)
library(DBI)
library(dbplyr)
library(dplyr)
library(bigrquery)
library(tidyverse)
library(plotly)
library(DT)
library(ggvis)
library(shiny)
library(gapminder)
library(RSocrata)
library(lubridate)
```

## Google BigQuery Project ID

I used Google BigQuery with the following project ID to initialize the crime data.

```{r}
project <- "planar-ember-256702"
```

## Connecting to Chicago Crime Database

I used the Chicago crime database and set up the connection to the database.

```{r}
con <- dbConnect(
  bigrquery::bigquery(),
  project = "bigquery-public-data",
  dataset = "chicago_crime",
  billing = project
)
con 
```

## Preparing Data

I wrote queries from within R via the `DBI` package. I am interested in the number of arrests grouped by year, district, and primary type of crime.

```{r}
#query object that fulfills the above
sql <- "SELECT year, district, primary_type, COUNT(*) AS arrests 
        FROM `bigquery-public-data.chicago_crime.crime` 
        WHERE arrest = TRUE
        GROUP BY year, district, primary_type"

#converting data to .csv to import into github
write.csv(crime,'C:\\Users\\Rebecca Oh\\Documents\\crimeapp\\crime.csv', row.names = FALSE)

#reading data from github
crime <- read_csv("https://raw.githubusercontent.com/daraeoh/Chicago-Crime-App/master/crime.csv")
```

#Data Cleaning
```{r}
crime$district[crime$district == "1"] <- "1 - Central"
crime$district[crime$district == "2"] <- "2 - Wentworth"
crime$district[crime$district == "3"] <- "3 Grand Crossing"
crime$district[crime$district == "4"] <- "4 - South Chicago"
crime$district[crime$district == "5"] <- "5 - Calumet"
crime$district[crime$district == "6"] <- "6 - Gresham"
crime$district[crime$district == "7"] <- "7 - Englewood"
crime$district[crime$district == "8"] <- "8 - Chicago Lawn"
crime$district[crime$district == "9"] <- "9 - Deering"
crime$district[crime$district == "10"] <- "10 - Ogden"
crime$district[crime$district == "11"] <- "11 - Harrison"
crime$district[crime$district == "12"] <- "12 - Near West"
crime$district[crime$district == "14"] <- "14 - Shakespeare"
crime$district[crime$district == "15"] <- "15 - Austin"
crime$district[crime$district == "16"] <- "16 - Jefferson Park"
crime$district[crime$district == "17"] <- "17 - Albany Park"
crime$district[crime$district == "18"] <- "18 - Near North"
crime$district[crime$district == "19"] <- "19 - Town Hall"
crime$district[crime$district == "20"] <- "20 - Lincoln"
crime$district[crime$district == "22"] <- "22 - Morgan Park"
crime$district[crime$district == "24"] <- "24 - Rogers Park"
crime$district[crime$district == "25"] <- "25 - Grand Central"
```

# Define UI
```{r}
ui <- fluidPage(
  
  # App title
  titlePanel("Chicago Crime Visualization"),
  
  br(),
  
  h5("Author: Rebecca Oh"),
  
  br(),
  
  # Sidebar layout with input and output definitions
  ui <- fluidPage(
  
   sidebarLayout(
     
     # Inputs
     sidebarPanel(
       
       # Select district number
       selectInput(inputId = "district",
                   label = "District",
                   choices = c("1 - Central", "2 - Wentworth", "3 - Grand Crossing", "4 - South Chicago", "5 - Calumet",
                               "6 - Gresham", "7 - Englewood", "8 - Chicago Lawn", "9 - Deering", "10 - Ogden", 
                               "11 - Harrison", "12 - Near West", "14 - Shakespeare", "15 - Austin", 
                               "16 - Jefferson Park", "17 - Albany Park", "18 - Near North", "19 - Town Hall", 
                               "20 - Lincoln", "21", "22 - Morgan Park", "24 - Rogers Park", 
                               "25 - Grand Central", "31") %>% unique(),
                   selected = "1 - Central"),
       
       # Slide year range
       sliderInput(inputId = "year",
                   label = "Range of years",
                   min = 2001,
                   max = 2019,
                   value = c(2001, 2019), 
                   sep = ""),
       
       # Select crime type
       checkboxGroupInput(inputId = "primary_type",
                     label = "Choose a primary type of crime",
                     choices = c("ARSON", "ASSAULT", "BATTERY", "BURGLARY", "CONCEALED CARRY LICENSE VIOLATION", 
                                 "CRIMINAL DAMAGE", "CRIMINAL TRESPASS", "CRIM SEXUAL ASSAULT", "DECEPTIVE PRACTICE", 
                                 "GAMBLING", "HOMICIDE", "INTERFERENCE WITH PUBLIC OFFICER", "INTIMIDATION", 
                                 "KIDNAPPING", "LIQUOR LAW VIOLATION", "MOTOR VEHICLE THEFT", "NARCOTICS", 
                                 "NON-CRIMINAL (SUBJECT SPECIFIED)", "OBSCENITY", "OTHER NARCOTIC VIOLATION", 
                                 "OTHER OFFENSE", "OFFENSE INVOLVING CHILDREN", "PROSTITUTION", "PUBLIC INDECENCY", 
                                 "PUBLIC PEACE VIOLATION", "ROBBERY", "SEX OFFENSE", "STALKING", "THEFT", "WEAPONS VIOLATION"),
                     selected = "ARSON")
     ),
     
     # Outputs
     mainPanel(
       position = "left",
       plotlyOutput(outputId = "scatterplot")
     )
   )
  )
)
```

# Server
```{r}
# Define server function
server <- function(input, output) {
  
  # Create data
  crime2 <- reactive({
    crime %>%
      filter(district == as.vector(input$district)) %>%
      filter(year >= as.integer(input$year[1]) & year <= as.integer(input$year[2])) %>%
      filter(primary_type == input$primary_type)
  })
  
  # Create scatterplot object
  output$scatterplot <- renderPlotly({
    p <- ggplot(data = crime2(), aes(x = year, y = arrests)) +
      geom_point(aes(color = primary_type)) +
      geom_smooth(method = 'loess') +
      theme_light() +
      labs(title = paste0("Chicago Crimes in District ", input$district),
           x = "Year",
           y = "Number of Arrests",
           color = "Primary Type of Crime")
    ggplotly(p)
  })
}
```

# Run the Shiny application
```{r}
shinyApp(ui = ui, server = server)
```